---
title: "CaseStudy1"
author: "Cory Nichols"
date: "2/28/2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

####Introduction
We are investigating GDP quantities by country. 

```{r importData, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(plyr)
library(ggplot2)

gdpURL <- 'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv'
eduURL <- 'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv'

gdp <- read.csv(gdpURL, skip = 3, stringsAsFactors = F)[2:191,c(1,2,4,5)] 
# excludes lines with nulls and summary data points (192:232)
edu <- read.csv(eduURL, stringsAsFactors = F)
```

####Data Cleaning and Setup

To clean the dataset, we'll rename the columns of the gdp dataset and transform gdp quantities to numeric

We will also join the gdp and education datasets on the code variable

```{r clean, echo=FALSE}
names(gdp) = c('code','rank','economy','usd')
gdp$usd <- as.numeric(gsub(",","",gdp$usd))
gdp$rank <- as.integer(gdp$rank)
names(edu)[1] = "code"
df <- plyr::join(gdp, edu, by = "code", type="inner")
nrow(df) # matched IDs
```

####Ordering GDP

In order to determine the 13th lowest GDP country in the dataset, we arrange and index the now-joined dataset:

```{r test}
df <- plyr::arrange(df, usd)
df[13,]['Short.Name'] # country name, 13th position
```


####Mean GDP Ranking by Category

We are specifically interested in determining the mean GDP rank position for High Income groupings. In order to investigate this datapoint, we transform rank

```{r AvgRank, echo = FALSE}
ddply(df, ~Income.Group, summarize, avgRank = mean(rank))
```

####Plotting GDP

Plot GDP positions

```{r plotter, echo= FALSE}
ggplot(data=df, aes(x=df$rank, y=df$usd, color=df$Income.Group))+geom_point(position="jitter")
```

####Lower Middle Income in Quantile 1

Detail here
```{r quantiles, echo = FALSE, message=FALSE}
#quantRank <- quantile(df$rank, c(.2, .4, .6, .8))
#cut(df$rank, 5, labels = c(1,2,3,4,5))
df$quantile <- cut(df$rank, 5, labels = c(1,2,3,4,5)) # why do we need quantiles here?
top38 <- subset(df, df$rank <= 38)
nrow(subset(top38, Income.Group == 'Lower middle income'))
```

####Conclusion

This is the conclusion